---
title: "Homework 1"
author: "Leon McSwain"
date: "2/20/2021"
output: html_document
---

```{r setup, include=FALSE}
library(R2jags)
HIV <- read.csv("C:/Users/12298/Desktop/BIOS560R/Assignment_One/HIV.csv") %>% as.data.frame()
```


Poisson and Exponential Analysis
```{r}
jags.model1 <- function(){
#Dist for data likelyhood
#looping through vector of case numbers for data likelyhood
  for (i in 1:N0) {
    C0[i] ~ dpois(lambda0)
    } #No Intervention
  for (i in 1:N1) {
    C1[i] ~ dpois(lambda1)
    } #Intervention
  
#Priors for mean cases - using exponential conjugate prior for fewest assumptions about case means - using 1/mean of all cases as the lambda for exp distribution
  lambda0 ~ dexp(P0) #Fewest assumptions with only one parameter
  lambda1 ~ dexp(P0) 
  
#Parameter to save - difference in mean case numbers as measure of risk reduction
  Risk=lambda0-lambda1
  
}

#Variables
C0 <- as.numeric((HIV$cases[HIV$status==0])) #Vector of non-treatment group
C1 <- as.numeric((HIV$cases[HIV$status==1])) #Vector of treatment group
N0 <- sum(HIV$status==0) #total cases for non-treatment group
N1 <- sum(HIV$status==1) #total cases for treatment group
P0=1/mean(HIV$cases) 

```


Running the Jags model. Saving lambda distributions (posterior means) and Risk parameter
```{r}
jags.dat=list(C0=C0, C1=C1, N0 = N0, N1 = N1, P0=P0)

fit = jags (data = jags.dat, model.file = jags.model1, parameters.to.save = c("Risk", "lambda0", "lambda1"), n.chain = 1, n.iter = 51000, n.burnin = 1000, n.thin = 1)
fit.mcmc <- as.mcmc(fit)

summary (fit.mcmc)
```

Analysis of posterior distribution results
```{r}
#Point and Interval Estimates
median(as.matrix(fit.mcmc[,4]))
HPDI <- HPDinterval(fit.mcmc[,4], credMass=0.89)

#Histograms of parameters and risk
#Non-treatment Group Posterior
hist(as.matrix(fit.mcmc[,2]), main="Non-intervention Group Posterior", xlab="Posterior Mean")
abline(v=mean(as.matrix(fit.mcmc[,2])))

#Treatment Group Posterior with Highest Posterior Density Interval
hist(as.matrix(fit.mcmc[,3]), main="Intervention Group Posterior", xlab="Posterior Mean") 
abline(v=mean(as.matrix(fit.mcmc[,3])))

#Risk Posterior 
hist(as.matrix(fit.mcmc[,4]), main="Risk Reduction", xlab="Mean Difference") #The risk = lambda1-Lambda0
abline(v=c(HPDI[[c(1,1)]], mean(as.matrix(fit.mcmc[,4])), HPDI[[c(1,2)]]), col=c("blue", "red", "blue"))

#Showing the percent difference between the two parameters to show the magnitude of the difference.
hist(((rate_prop - 1)*100), main="Percent Risk Difference Between Groups", xlab="Percent of Mean Proportion")
abline(v=c(mean(rate_prop-1)*100))

# Posterior probability that control > treatment
sum(rate_prop > 1) /length(rate_prop)


```



Negative Binomial and Beta Analysis
Important questions: 
Would the number of failures before success be represented as a minimum difference in the mean case number? 
###########################################################################################################################
```{r}
jags.model1 <- function(){
#Dist for data likelihood
#looping through vector of case numbers for data likelihood
  for (i in 1:N0) {
    C0[i] ~ dnegbin(p0, r0)
    } #No Intervention
  for (i in 1:N1) {
    C1[i] ~ dnegbin(p1, r1)
    } #Intervention
  
#Re-paramaterizing in terms of mean
  p0 <- r0/(r0+m0)
  p1 <- r1/(r1+m1)
  
#Priors for mean cases (probability parameter)
  m0 ~ dexp(P0)
  m1 ~ dexp(P0)
  
#Priors for variance - Dispersion parameter
  r0 ~ dexp(1)
  r1 ~ dexp(1)
  
#Parameter to save - difference in mean case numbers as measure of risk reduction
  Risk=m0-m1
  
  var0 = r0*(1-p0)/(p0*p0)
  var1 = r1*(1-p1)/(p1*p1)
  
}

#Variables
C0 <- as.numeric((HIV$cases[HIV$status==0])) #Vector of non-treatment group
C1 <- as.numeric((HIV$cases[HIV$status==1])) #Vector of treatment group
N0 <- sum(HIV$status==0) #total cases for non-treatment group
N1 <- sum(HIV$status==1) #total cases for treatment group
P0=1/mean(HIV$cases) 

```


Running the Jags model. Saving lambda distributions (posterior means) and Risk parameter
```{r}
jags.dat=list(C0=C0, C1=C1, N0 = N0, N1 = N1, P0=P0)

fit = jags (data = jags.dat, model.file = jags.model1, parameters.to.save = c("Risk", "m0", "m1", "var1", "var0"), n.chain = 1, n.iter = 51000, n.burnin = 1000, n.thin = 1)
fit.mcmc <- as.mcmc(fit)

summary (fit.mcmc)
```

```{r}
#Point and Interval Estimates
median(as.matrix(fit.mcmc[,4]))
HPDI <- HPDinterval(fit.mcmc[,4], credMass=0.89)

#Histograms of parameters and risk
#Non-treatment Group Posterior
hist(as.matrix(fit.mcmc[,2]), main="Non-intervention Group Posterior", xlab="Posterior Mean")
abline(v=mean(as.matrix(fit.mcmc[,2])))

#Treatment Group Posterior with Highest Posterior Density Interval
hist(as.matrix(fit.mcmc[,3]), main="Intervention Group Posterior", xlab="Posterior Mean") 
abline(v=mean(as.matrix(fit.mcmc[,3])))

#Risk Posterior 
hist(as.matrix(fit.mcmc[,4]), main="Risk Reduction", xlab="Mean Difference") 
abline(v=c(HPDI[[c(1,1)]], mean(as.matrix(fit.mcmc[,4])), HPDI[[c(1,2)]]), col=c("blue", "red", "blue"))

#Showing the percent difference between the two parameters to show the magnitude of the difference.
rate_prop <- as.matrix(fit.mcmc[,2])/as.matrix(fit.mcmc[,3])
hist(((rate_prop - 1)*100), main="Percent Risk Difference Between Groups", xlab="Percent of Mean Proportion")
abline(v=c(mean(rate_prop-1)*100))

# Posterior probability that control > treatment
sum(rate_prop > 1) /length(rate_prop)


```


